---
title: Using Bioconductor with High Throughput Assays
output:
  rmdformats::readthedown:
    toc_depth: 3
    number_sections: true
---

Bioconductor includes packages for analysis of diverse areas of
high-throughput assays such as flow cytometry, quantitative real-time PCR,
mass spectrometry, proteomics and other cell-based data.

# Sample Workflow

The following psuedo-code illustrates a typical R / Bioconductor
session. It makes use of the flow cytometry packages to load, transform and
visualize the flow data and gate certain populations in the dataset. 

The workflow loads the `flowCore`, `flowStats` and `flowViz` packages and its
dependencies.  It loads the ITN data with 15 samples, each of which includes,
in addition to FSC and SSC, 5 fluorescence channels: CD3, CD4, CD8, CD69 and
HLADR. 

```{r echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages(library(flowCore))
suppressPackageStartupMessages(library(flowStats))
suppressPackageStartupMessages(library(flowViz))
```

```{r}
## Load packages
library(flowCore)
library(flowStats)
library(flowViz) # for flow data visualization

## Load data
data(ITN)
ITN
```

First, we need to transform all the fluorescence channels. Using a `workFlow`
object can help to keep track of our progress.

```{r}
## Create a workflow instance and transform data using asinh
wf <- workFlow(ITN)
asinh <- arcsinhTransform()
tl <- transformList(colnames(ITN)[3:7], asinh, 
                      transformationId = "asinh")
add(wf, tl)
```

Next we use the `lymphGate` function to find the T-cells in the CD3/SSC
projection.
    
```{r}
## Identify T-cells population
lg <- lymphGate(Data(wf[["asinh"]]), channels=c("SSC", "CD3"),
         preselection="CD4", filterId="TCells", eval=FALSE,
         scale=2.5)
add(wf, lg$n2gate, parent="asinh")
print(xyplot(SSC ~ CD3| PatientID, wf[["TCells+"]],
             par.settings=list(gate=list(col="red", 
             fill="red", alpha=0.3))))
```

    
A typical workflow for flow cytometry data analysis in Bioconductor flow
packages include data transformation, normalization, filtering, manual gating,
semi-automatic gating and automatic clustering if desired. Details can be
found in [flowWorkFlow.pdf](flowWorkFlow.pdf) or the vignettes of the
[flow cytometry packages](#diverse-assays-resources).

# Installation and Use

Follow [installation instructions](/install/) to start using these
packages.  To install the `flowCore` package and all of its
dependencies, evaluate the commands

```{r eval=FALSE}
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite("flowCore")
```

Package installation is required only once per R installation. View a
full list of
[available packages](/packages/release/bioc/).

To use the `flowCore` package, evaluate the command

```{r eval=FALSE}
library("flowCore")
```

This instruction is required once in each R session.

# Exploring Package Content

Packages have extensive help pages, and include vignettes highlighting
common use cases. The help pages and vignettes are available from
within R. After loading a package, use syntax like

    help(package="flowCore")
    ?read.FCS

to obtain an overview of help on the `flowCore` package, and the
`read.FCS` function, and

```{r eval=FALSE}
browseVignettes(package="flowCore")
```

to view vignettes (providing a more comprehensive introduction to
package functionality) in the `flowCore` package. Use

```{r eval=FALSE}
help.start()
```
to open a web page containing comprehensive help resources.

# Diverse Assays Resources

The following provide a brief overview of packages useful for analysis
of high-throughput assays. More comprehensive workflows can be found
in documentation (available from [package
descriptions](/packages/release/bioc/))
and in Bioconductor [publications](/help/publications/).

## Flow Cytometry ##

These packages use standard FCS files, including infrastructure,
utilities, visualization and semi-autogating methods for the
analysis of flow cytometry data.

[flowCore](/packages/release/bioc/html/flowCore.html),
[flowViz](/packages/release/bioc/html/flowViz.html),
[flowQ](/packages/release/bioc/html/flowQ.html),
[flowStats](/packages/release/bioc/html/flowStats.html),
[flowUtils](/packages/release/bioc/html/flowUtils.html),
[flowFP](/packages/release/bioc/html/flowFP.html),
[flowTrans](/packages/release/bioc/html/flowTrans.html),

Algorithms for clustering flow cytometry data are found in these packages:

[flowClust](/packages/release/bioc/html/flowClust.html),
[flowMeans](/packages/release/bioc/html/flowMeans.html),
[flowMerge](/packages/release/bioc/html/flowMerge.html),
[SamSPECTRAL](/packages/release/bioc/html/SamSPECTRAL.html)

A typical workflow using the packages `flowCore`, `flowViz`, `flowQ` and
`flowStats` is described in detail in [flowWorkFlow.pdf](flowWorkFlow.pdf).
The data files used in the workflow can be downloaded from
[here](dataFiles.tar).

## Cell-based Assays ##

These packages provide data structures and algorithms for cell-based
high-throughput screens (HTS).

[cellHTS2](/packages/release/bioc/html/cellHTS2.html),
[RNAither](/packages/release/bioc/html/RNAither.html)

This package supports the xCELLigence system which contains a series of
real-time cell analyzer (RTCA).

[RTCA](/packages/release/bioc/html/RTCA.html)

## High-throughput qPCR Assays ##

These package provide algorithm for the analysis of cycle threshold
(Ct) from quantitative real-time PCR data.

[HTqPCR](/packages/release/bioc/html/HTqPCR.html),
[ddCt](/packages/release/bioc/html/ddCt.html),
[qpcrNorm](/packages/release/bioc/html/qpcrNorm.html)

## Mass Spectrometry and Proteomics data ##

These packages provide framework for processing, visualization, and
statistical analysis of mass spectral and proteomics data.

[clippda](/packages/release/bioc/html/clippda.html),
[MassArray](/packages/release/bioc/html/MassArray.html),
[MassSpecWavelet](/packages/release/bioc/html/MassSpecWavelet.html),
[PROcess](/packages/release/bioc/html/PROcess.html),
[flagme](/packages/release/bioc/html/flagme.html),
[xcms](/packages/release/bioc/html/xcms.html)

## Imaging Based Assays ##

These packages provide infrastructure for image-based phenotyping and automation of other image-related tasks:

[EBImage](/packages/release/bioc/html/EBImage.html),
[imageHTS](/packages/release/bioc/html/imagHTS.html)

```{r}
sessionInfo()
```
